---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# Drilling Data (2018–2024)

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)

<!-- badges: end -->

This dataset captures detailed information from water point drilling projects conducted between 2018 and 2024. The data was collected from various borehole sites and includes a wide range of attributes covering the full lifecycle of a drilling project — from planning and site setup to borehole development and evaluation.

Key Components:
Administrative & Contextual Info: Includes the date of drilling, GPS coordinates, purpose of the water point, funding source, and responsible contractors.

Pre-Drilling Activities: Captures site selection methodologies, availability of government supervisors, community agreements, hydrogeological studies, and geophysical survey results.

Site & Equipment Setup: Details on rig setup, equipment condition, safety features, PPE availability, site accessibility, and distance from potential contamination sources.

Drilling Operations: Documents the type of rig used, compressor specs, drill rod length, drilling method, bit type, borehole diameter, depth intervals, and drilling time.

Hydrogeological Observations: Tracks water strikes, rock/soil formations, grain size, sorting, weathering, and additional geological descriptions.

Borehole Development & Testing: Includes discharge rates, EC, TDS, pH, temperature, turbidity, sediment, and static water levels (pre and post development).

Casing & Installation Details: Information on casing materials, diameters, slot aperture, gravel pack quality, and final borehole design.

Safety & Environmental Checks: Evaluates welfare facilities, fire extinguishers, first aid kits, equipment protection, and chemical storage practices.

Multimedia Links: Some entries include links to images from the drilling sites hosted on the mWater platform.

Example Use Cases:
Analyzing factors that affect borehole success rates.

Tracking contractor compliance with safety and environmental standards.

Evaluating relationships between hydro-geological conditions and borehole performance.

Enhancing water resource planning and monitoring in rural development contexts.

Note: This dataset contains blank or missing values in some fields, especially where data collection was not applicable. Users should apply appropriate data cleaning and preprocessing methods before analysis.



## Installation

You can install the development version of drillingdata from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/drillingdata")
```

```{r}
## Run the following code in console if you don't have the packages
## install.packages(c("dplyr", "knitr", "readr", "stringr", "gt", "kableExtra"))
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(gt)
library(kableExtra)
```

Alternatively, you can download the individual datasets as a CSV or XLSX
file from the table below.

1.  Click Download CSV. A window opens that displays the CSV in
    your browser.
2.  Right-click anywhere inside the window and select "Save Page As...".
3.  Save the file in a folder of your choice.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

extdata_path <- "https://github.com/openwashdata/drillingdata/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  dplyr::mutate(file_name = str_remove(file_name, ".rda")) |> 
  dplyr::rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

## Data
The package provides access to the Drilling data survey conducted between 2018 and 2024

```{r}
library(drillingdata)
```

### drillingdata

The dataset `drillingdata` has `r nrow(drillingdata)` observations and `r ncol(drillingdata)`
variables

```{r}
drillingdata |> 
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "drillingdata.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

## Example

```{r}
library(drillingdata)

# Example code to display a bar graph for the boreholes drilled per year
library(ggplot2)
library(lubridate)
library(dplyr)

# Convert 'date_of_drilling' to Date format and extract the year
drillingdata$year <- year(mdy(drillingdata$date_of_drilling))

# Count number of boreholes drilled per year
boreholes_per_year <- drillingdata %>%
  filter(!is.na(year)) %>%
  group_by(year) %>%
  summarise(boreholes_drilled = n())

# Create the bar plot
ggplot(boreholes_per_year, aes(x = factor(year), y = boreholes_drilled)) +
  geom_col(fill = "red") +
  theme_minimal() +
  labs(
    title = "Boreholes Drilled Per Year",
    x = "Year",
    y = "Number of Boreholes Drilled"
  )

```

## License

Data are available as
[CC-BY](https://github.com/openwashdata/%7B%7B%7Bpackagename%7D%7D%7D/blob/main/LICENSE.md).

## Citation

Please cite this package using:

```{r}
citation("drillingdata")
```
